Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-07-03 text/paper/paper.tex
paper reworked motivation section, more typicality

\usepackage[final,            % override "draft" which means "no nothing"
            colorlinks,       % rather than outlining them in boxes
            linkcolor=gray,   % override truly awful colour choices
            citecolor=gray,   %   (ditto)
            urlcolor=gray,    %   (ditto)
            plainpages=false, % to overcome complaints with multiple
            pdfpagelabels,    % multiple page 1-s due to preface
            hypertexnames=false % solves warning, but interferes with
                                % index and \autoref apparently
            ]{hyperref}

\usepackage{amsmath}            % Formeln
\usepackage{amsfonts}           % Fonts for Formulas
\usepackage{amssymb}
\usepackage[final]{graphicx}
\usepackage{booktabs}
\usepackage{enumerate}
\usepackage{lipsum}
\usepackage{txfonts} % for strict implication symbols
\usepackage{soul}
\usepackage{relsize} % provides command \relsize{+/-x} for relative
                     % font size changes
\usepackage[german,english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} 
\usepackage{subfig}
\usepackage{xypic}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}
\usepackage{refcount}
\usepackage{xspace} % for \xspace in definition of acronyms etc.
\usepackage{attrib} % for right-aligned references at the end of
                    % quotes; part of Frankenstein bundle, but 
\renewcommand\PreTrib {}   % overwrite these commands, from attrib.sty
\renewcommand\PostTrib {}  % to suppress additional brackets around
                           % attributions

\usepackage{pgfplots}

@@

\citeauthor{ChemlaSpector2010:Experimental-Ev} hypothesized that the
degree to which a sentence was rated acceptable is proportional to the
number of available true readings. Observed averaged clicking
positions are shown in Figure~\ref{fig:Chemla-Spector-Results}.
%
\begin{figure}
  \centering
  \subfloat[][\as-condition]{
    \label{fig:CS-Results-as}
        \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_AE.png}
  }
  \subfloat[test][\es-condition]{
    \label{fig:CS-Results-es}
    \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_GE.png}
  }
  \caption{Results from
    \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
    (\citeyear{ChemlaSpector2010:Experimental-Ev}) study}
  \label{fig:Chemla-Spector-Results}
\end{figure}
%
According to \citeauthor{ChemlaSpector2010:Experimental-Ev}, the
crucial piece of evidence for the availability of local readings for
\as-sentences was that these sentences yielded higher graded
acceptability scores for the strong situation than for the weak
situation (although these differ only with respect to the truth value
of the local reading). In a similar way, evidence for the availability
of the local reading for \es-sentences comes from the difference
between the local and the literal situation. Strikingly, in the latter
case, the \es-sentences received an average of 73\% acceptability in
the local situation although the literal and global readings are false
in this case.

\subsection{Reflection}
\label{sec:critique}

Summing up, the mentioned studies have presented diverging evidence
concerning the availability of local readings. Various problematic
aspects of these studies have already been discussed elsewhere
\citep[e.g.][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,Tielvan-Tiel2012:Embedded-Scalar}. Here,
we want to point out two particular methodological problems that do
impede their interpretation. However, we also want to make clear that
these problems follow almost directly from the nature of the
phenomenon under investigation and therefore seem nearly impossible to
avoid.

Firstly, a mapping between responses in a picture-verification task
and readings of the relevant sentences is problematic due to the
logical dependencies between the relevant readings discussed in
Section~\ref{sec:get-know-your}. Let us consider \as-sentences in the
following because these have been addressed by all three
aforementioned studies. (Similar arguments apply to \es-sentences.) It
is impossible to construct situations that clearly correspond to local
readings of \as-sentences, since the local reading of such sentences
implies their global and local readings. Every situation which is
consistent with such a sentence under its local reading is also
consistent with the other two readings. Therefore, we cannot test
directly and independently all the cases we might wish to using a
simple picture-verification paradigm. It would be especially desirable
to test situations that are consistent with the local reading, but
inconsistent with the global and literal reading. In that case
\emph{yes}-responses would unambiguously indicate availability of
local readings. However, since this is impossible we are left with
suboptimal choices. In the experiment of
\citet{GeurtsPouscoulous2009:Embedded-Implic} situations that are
inconsistent with local readings but consistent with the other two
readings were tested. In that case \emph{no}-responses would be decisive~--
when compared to a suitable baseline. However, \emph{yes}-responses were
observed exclusively. These can be attributed to one of the other
readings. Therefore, assuming a principle of charity
\citep{Davidson1973:Radical-Interpr} to be at play in picture
verification tasks, the data of Geurts do not provide evidence against
the existence of local readings.  In the experiments of C\&S and C\&D
on the other hand situations that are consistent with all three
readings were tested. Here again 'yes'-judgments cannot be interpreted
confidently because they can always be attributed to global or literal
readings. Both C\&S and C\&D attempt to surpass this problem by
interpreting relative preferences. This does, however, bring us to the
second problem.

Preferences may be affectd by graphical properties of the picture
materials. For example, verification complexity or typicality of the
pictures with respect to the test sentences may differ between
conditions and thereby induce certain preferences. In his careful
critique van Tiel has eleborated on this point. Van Tiel points out
that the relative preferences C\&S and C\&D interpret can be explained
solely in terms of typicality, because the `strong'-situation (see
e.g.~Figure~\ref{fig:strong}) are more typical instances of the
sentence meaning than the competing pictures, irrespective of local
implicatures.

To conclude, in order to decide about the availablity of local
implicatures we would ideally like to test situations that can
unambiguously be mapped to the different readings and do not differ in
typcality or other crucial properties. Given the logical dependencies
inherent to the relevant readings and the considerations concerning
typicality presented by van Tiel this seems impossible to achieve. As
we willl argue below we do, however, believe it is.

@@

\begin{itemize}
\item K1:
  \begin{itemize}
  \item tried to replicate
    \citet{GeurtsPouscoulous2009:Embedded-Implic}'s study
  \item but tested all conditions
  \item used German cleft-sentences (possibly problematic)
  \item some evidence for subjective patterns in answers
  \end{itemize}
\item We also tested examples like (\ref{bsp:EA}), which are the ``global
alternatives'' to \es-sentences, with a universal quantifier where
\es-sentences have scalar \emph{some}.
\begin{exe}
\ex \gll F\"{u}r \mymark{genau} \mymark{eines} dieser Kinder gilt: es
  mag \mymark{jede} dieser Speisen. \label{bsp:EA}\\
  For exactly one {of these} children {it holds} it likes all {of
    these} dishes.\\
  For exactly one of these children it holds that it likes all of these dishes.
\end{exe}
Sentence like (\ref{bsp:EA}) are literally true in both of the
situations given in Figure~\ref{fig:EA-distinguishing-pics}, but these
situations differ in the truth value of the corresponding
\es-sentence. Interestingly, we found a difference in acceptability,
but a curious one. The sentence was judged \emph{true} in 90.34\% of
the cases for the situation in Figure~\ref{fig:global-EA} and in
98.34\% in situation \ref{fig:literal-EA}. The observed difference is
significant ($t = -3.8259$, $df = 322.375$, $p < 0.001$). That is a
curious finding, because, from the point of view of implicature
calculation, the only thing that could plausibly explain this
acceptability difference is the presence of a global implicature,
resulting from the negation of the corresponding \es-sentence. But
such a putative global implicature would be true in the
situation~\ref{fig:global-EA}, but false in the
situation~\ref{fig:literal-EA}. Consequently, the observed difference
in acceptability cannot be explained simply in terms of the hypothesis
of \citet{ChemlaSpector2010:Experimental-Ev} that acceptability is
proportional to the number of available true readings: if a putative
global reading is at stake, the acceptability pattern should be
reversed; if no such global reading is at stake, then the
acceptability difference cannot be explained.
\end{itemize}

>>>>

\usepackage[final,            % override "draft" which means "no nothing"
            colorlinks,       % rather than outlining them in boxes
            linkcolor=gray,   % override truly awful color choices
            citecolor=gray,   %   (ditto)
            urlcolor=gray,    %   (ditto)
            plainpages=false, % to overcome complaints with multiple
            pdfpagelabels,    % multiple page 1-s due to preface
            hypertexnames=false % solves warning, but interferes with
                                % index and \autoref apparently
            ]{hyperref}

\usepackage{amsmath}            % Formeln
\usepackage{amsfonts}           % Fonts for Formulas
\usepackage{amssymb}
\usepackage[final]{graphicx}
\usepackage{booktabs}
\usepackage{enumerate}
\usepackage{lipsum}
\usepackage{txfonts} % for strict implication symbols
\usepackage{soul}
\usepackage{relsize} % provides command \relsize{+/-x} for relative
                     % font size changes
\usepackage[german,english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} 
\usepackage{subfig}
\usepackage{xypic}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}
\usepackage{refcount}
\usepackage{xspace} % for \xspace in definition of acronyms etc.
\usepackage{attrib} % for right-aligned references at the end of
                    % quotes; part of Frankenstein bundle, but 
\renewcommand\PreTrib {}   % overwrite these commands, from attrib.sty
\renewcommand\PostTrib {}  % to suppress additional brackets around
                           % attributions
\usepackage{eurosym}
\usepackage{pgfplots}

@@

\citeauthor{ChemlaSpector2010:Experimental-Ev} hypothesized that the
degree to which a sentence was rated acceptable is proportional to the
number of available true readings. Observed averaged clicking
positions are shown in Figure~\ref{fig:Chemla-Spector-Results}.
%
\begin{figure}
  \centering
  \subfloat[][\as-condition]{
    \label{fig:CS-Results-as}
        \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_AE.png}
  }
  \subfloat[test][\es-condition]{
    \label{fig:CS-Results-es}
    \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_GE.png}
  }
  \caption{Results from
    \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
    (\citeyear{ChemlaSpector2010:Experimental-Ev}) study}
  \label{fig:Chemla-Spector-Results}
\end{figure}
%
According to \citeauthor{ChemlaSpector2010:Experimental-Ev}, the
crucial piece of evidence for the availability of local readings for
\as-sentences is that these sentences yielded higher graded
acceptability scores for the strong situation than for the weak
situation (although these differ only with respect to the truth value
of the local reading). In a similar way, evidence for the availability
of the local reading for \es-sentences comes from the difference
between the local and the literal situation. Strikingly, in the latter
case, the \es-sentences received an average of 73\% acceptability in
the local situation although the literal and global readings are false
in this case.

\subsection{Reflection: Categorical Judgements \& Typicality}
\label{sec:local-read-categ}

Summing up so far, three previous studies have presented different
pieces of evidence concerning putative local readings. G\&P's results
suggest that local readings are not available, because they do not
affect a categorical truth-value judgments in a picture-verification
task. On the other hand, C\&D's and C\&S's design was suggested to be
sensitive enough to reveal effects of local readings: either by a
choice of the situation that better befits a given sentence (C\&D), or
by a graded truth-value judgment (C\&S).

To account for all of the available data in conjunction, it could be
hypothesized along with C\&D and C\&S that local readings are
available, but so dispreferred they do not affect simple
\emph{categorical} truth-value judgements. This, however, does not
seem to be the case, as revealed by a pilot study of ourselves. We
tried to replicate G\&P's results, using a simple categorial
truth-value judgement task on pictures that were exactly like those in
Figures \ref{fig:AS-distinguishing-pics} and
\ref{fig:ES-distinguishing-pics}. Test sentences were German
cleft-constructions as in (\ref{bsp:AS-K1}) and
(\ref{bsp:ES-K1}).\fn{We concede that these cleft-constructions might
  unduly emphasize potential local readings, but that is not
  essential, because the main issue we are trying to address here is
  whether putative local readings \emph{could} affect categorical
  truth-value judgements.}
\begin{exe}
\ex \gll F\"{u}r \mymark{jedes} dieser Kinder gilt: es
  mag \mymark{einige} dieser Speisen. \label{bsp:AS-K1}\\
  For every {of these} children {it holds} it likes some {of
    these} dishes.\\
  For all of these children it holds that it likes some of these dishes.
\end{exe}
\begin{exe}
\ex \gll F\"{u}r \mymark{genau} \mymark{eines} dieser Kinder gilt: es
  mag \mymark{einige} dieser Speisen. \label{bsp:ES-K1}\\
  For exactly one {of these} children {it holds} it likes some {of
    these} dishes.\\
  For exactly one of these children it holds that it likes some of these dishes.
\end{exe}
60 native German speakers (40 female, mean age: $24$, ranging between
$20$ and $40$) participated in this pilot. Participants were na\"ive
to the purpose of the study and were paid \euro{10}
compensation.\dn{We should expand on the method in an appendix}

Mean acceptance rates for our target
sentences were (see also Figure~\ref{fig:Bars-K1}):
\begin{center}
  \begin{tabular}{ccccccccc}
    \multicolumn{4}{c}{\as-sentences} &&
    \multicolumn{4}{c}{\es-sentences}     \\ \cmidrule(r){1-4} \cmidrule(r){6-9}
    false & literal & weak & strong && false & local & literal & all
    \\ \midrule 
    .9 & 72 & 79 & 90 && .8 & 20 & 71 & 93 
  \end{tabular}
\end{center}
Two observations are of interest here. Firstly, unlike G\&P, we
observed \emph{false}-judgements in the \as-weak condition, as well as
\emph{true}-judgements in the \es-local condition. Secondly, the
observed difference between the \as-weak and the \as-strong
conditions, on the one had, and the \es-false and the \es-local
condition, on the other hand, both reached significance.\dn{spell out
  in appendix}

\begin{figure}[]
\centering
\subfloat[\as-sentences]{
\includegraphics[width=5.5cm]{../pictures/paper/JudgmentsAllSomeBars.pdf}
\label{fig:EABars}
}
\subfloat[\es-sentences]{
\includegraphics[width=5cm]{../pictures/paper/JudgmentsExactlySomeBars.pdf}
\label{fig:EEBars}
}
\caption{Means of \emph{true}-judgments in our pilot study}
\label{fig:Bars-K1}
\end{figure}

The perhaps most obvious explanation for these observed patterns is
that local readings \emph{are} available and \emph{can} have an effect
even on categorical truth-value judgements. However, we believe that
this interpretation of the data is problematic for reasons already
addressed also by \citet{ChemlaSpector2010:Experimental-Ev} and, more
recently, \citet{Tielvan-Tiel2012:Embedded-Scalar}. The problem is
that the present variants of truth-value judgment tasks all run risk
of conflating evidence about the availability or preference of
readings with properties of the pictorial material that are
independent of the reading to be assessed. This problem affects all of
the studies covered above, to some degree or other, because it follows
almost directly from the nature of the phenomenon under investigation
and therefore seems nearly impossible to avoid.

To illustrate the potential conflation, consider a sentence like
(\ref{bsp:EA}), which is a ``global alternatives'' to an \es-sentence,
with a universal quantifier where an \es-sentence has scalar
\emph{some}.
\begin{exe}
\ex \gll F\"{u}r \mymark{genau} \mymark{eines} dieser Kinder gilt: es
  mag \mymark{jede} dieser Speisen. \label{bsp:EA}\\
  For exactly one {of these} children {it holds} it likes all {of
    these} dishes.\\
  For exactly one of these children it holds that it likes all of these dishes.
\end{exe}
Sentence like (\ref{bsp:EA}) are literally true in both of the
situations given in Figure~\ref{fig:EA-distinguishing-pics}, but these
situations differ in the truth value of the corresponding
\es-sentence. Our pilot study revealed a difference in acceptability,
but a curious one. The sentence was judged \emph{true} in 90.34\% of
the cases for the situation in Figure~\ref{fig:global-EA} and in
98.34\% in situation \ref{fig:literal-EA}. The observed difference is
significant ($t = -3.8259$, $df = 322.375$, $p < 0.001$). This is a
curious finding, because, from the point of view of implicature
calculation, the only thing that could plausibly explain this
acceptability difference is the presence of a global implicature,
resulting from the negation of the corresponding \es-sentence. But
such a putative global implicature would be true in the
situation~\ref{fig:global-EA}, but false in the
situation~\ref{fig:literal-EA}. Consequently, the observed difference
in acceptability cannot be explained simply in terms of the hypothesis
of \citet{ChemlaSpector2010:Experimental-Ev} that acceptability is
proportional to the number of available true readings. Rather some
feature of the pictorial material that is independent of the putative
readings under scrutiny does affect truth-value judgements, even
categorical ones.

\citet{ChemlaSpector2010:Experimental-Ev} acknowledge the possibility
that the \emph{typicality} of a picture may affect graded truth-value
judgements. Stronger still, they show that graded judgments of
\as-sentences differ significantly for different instantiations of the
conditions in \ref{fig:AS-distinguishing-pics}, that differ in the
amount of connections, but not so that they would alter the assignment
of truth-values to candidate
readings. \citeauthor{ChemlaSpector2010:Experimental-Ev} suggest that
typicality of the pictorial material can account for these
differences, but submit that this does not explain away the effect of
the local reading. This is disputed by
\citet{Tielvan-Tiel2012:Embedded-Scalar}, who demonstrates that all of
\citeauthor{CliftonDube2010:Embedded-Implic}'s and
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s data on \as-sentences
can be explained as typicality effects, dispensing the need to appeal
to any influence of local readings. We are sympathetic towards
\citeauthor{Tielvan-Tiel2012:Embedded-Scalar}'s innovative line of
reinterpretation of the data, but note, as
\citet{ChemlaSpector2010:Experimental-Ev} already observed, that his
typicality-based explanation does not extend to \es-sentences in an
obvious way\fn{For similar reasons, appeal to typicality cannot
  explain, as far as we can see, the
  preference for the situation in Figure~\ref{fig:literal-EA} over
  that in Figure\ref{fig:global-EA} given a sentence like
  (\ref{bsp:EA}).} and that it also does not explain why we observed putative
effects of typicality also in categorical judgements in our pilot.

To conclude, in order to decide about the availablity of local
implicatures we would ideally like to test situations that can
unambiguously be mapped to the different readings and do not differ in
typicality or other crucial properties. Given the logical dependencies
inherent to the relevant readings and the considerations concerning
typicality presented by van Tiel this seems impossible to achieve. As
we willl argue below we do, however, believe it is.

\begin{itemize}
\item what's needed
  \begin{itemize}
  \item evidence about all preferences
  \item factor out picture-complexity
  \item preferably categorical judgements
  \item intonation
  \end{itemize}

\end{itemize}

@@

\begin{itemize}
\item K1:
  \begin{itemize}
  \item tried to replicate
    \citet{GeurtsPouscoulous2009:Embedded-Implic}'s study
  \item but tested all conditions
  \item used German cleft-sentences (possibly problematic)
  \item some evidence for subjective patterns in answers
  \end{itemize}
\item Thirdly, we observed that a tendency that participants
who rejected \as-sentences in the literal situation where more prone
to also accept \es-sentences in the local condition.\dn{spell out
  in appendix} This suggests that there might be participants who are
more likely to be influenced by local readings than others.
\end{itemize}
